A Formal Practical Foundation of Bidirectional Typing for Simple Types

### 2023-06-10 tex/BiSig.tex
[ tex ] Fix some typos

\begin{theorem}[Decidability for Bidirectional Typing Judgements]
  For every type signature $\Sigma$ and bidirectional binding signature $\Omega$ for $\Sigma$ that is mode correct, the following two statements hold constructively.
  \begin{enumerate}
    \item For every context $\Gamma$ and a pre-synthesised term $\isTerm{t}$, it is decidable that there exists a type $A$ and a bidirectional typing derivation of
      \[
        \Gamma \vdash_{\Sigma, \Omega} \isTerm{t} \Rightarrow A.
      \]
    \item For every context $\Gamma$, a pre-checked term $\isTerm{t}$, and a type $A$, it is decidable that there is a typing derivation of
      \[
        \Gamma \vdash_{\Sigma, \Omega} \isTerm{t} \Leftarrow A.
      \]
  \end{enumerate}
\end{theorem}

@@

\begin{theorem}{Uniqueness of Synthesised Types}\label{thm:unique-syn}

@@

    \item For every set $\Xi$ of type variables, a list $\Delta$ of types with variables in $\Xi$, a partial substitution $\rho\colon \PSub{\Xi, \emptyset}$ defined for all types in $\Delta$, and a raw term with mode $d$ then 
      \begin{enumerate}
        \item if $d$ is $\Rightarrow$, then it is decidable that there exists a type $A$ and a typing derivation of
          \[
            \Gamma, \simsub{\Delta}{\rho} \vdash_{\Sigma, \Omega} t \Rightarrow A 
          \]
        \item if $d$ is $\Leftarrow$, then for every type $A$ it is decidable that there is a typing derivation of 
          \[
            \Gamma, \simsub{\Delta}{\rho} \vdash_{\Sigma, \Omega} t \Leftarrow A.
          \]
      \end{enumerate}
  \end{enumerate}
\end{theorem}
By mutual induction on raw terms $t$ with a mode. 
Moreover, if the type $A$ of $t$ is synthesised but checked against $B$, i.e.\ $t\uparrow$, but $A \neq B$, then \Cref{thm:unique-syn} is applied to derive a contradiction.
\begin{proof}
\end{proof}

>>>>

\begin{theorem}[Decidability for Bidirectional Typing Judgements]
  For every type signature $\Sigma$ and bidirectional binding signature $\Omega$ for $\Sigma$ that is mode correct, the following two statements hold constructively.
\LT{emphasise that LEM is not allowed.}
  \begin{enumerate}
    \item For every context $\Gamma$ and a pre-synthesised term $\isTerm{t}$, it is decidable that there exists a type $A$ and a bidirectional typing derivation of
      \[
        \Gamma \vdash_{\Sigma, \Omega} \isTerm{t} \Rightarrow A.
      \]
    \item For every context $\Gamma$, a pre-checked term $\isTerm{t}$, and a type $A$, it is decidable that there is a typing derivation of
      \[
        \Gamma \vdash_{\Sigma, \Omega} \isTerm{t} \Leftarrow A.
      \]
  \end{enumerate}
\end{theorem}

@@

\begin{theorem}[Uniqueness of Synthesised Types]\label{thm:unique-syn}

@@

    \item For every set $\Xi$ of type variables, a list $\Delta$ of types with variables in $\Xi$, a partial substitution $\rho\colon \PSub{\Xi, \emptyset}$ defined at least for all types in $\Delta$, a raw term $t$ with mode $d$, and a list of variables $\vec{x}$ then 
      \begin{enumerate}
        \item if $d$ is $\Rightarrow$, then it is decidable that there exists a type $A$ and a typing derivation of
          \[
            \Gamma, \vec{x} : \simsub{\Delta}{\rho} \vdash_{\Sigma, \Omega} t \Rightarrow A 
          \]
        \item if $d$ is $\Leftarrow$, then for every type $A$ it is decidable that there is a typing derivation of 
          \[
            \Gamma, \vec{x} : \simsub{\Delta}{\rho} \vdash_{\Sigma, \Omega} t \Leftarrow A.
          \]
      \end{enumerate}
  \end{enumerate}
\end{theorem}
By mutual induction on raw terms $t$ with a mode. 
Moreover, if the type $A$ of $t$ is synthesised but checked against $B$, i.e.\ $t\uparrow$, but $A \neq B$, then \Cref{thm:unique-syn} is applied to derive a contradiction.
\begin{proof}
\end{proof}
